Forecasting the old age dependency ratio to determine the best pension age

### 2019-01-04 PensionAge.Rmd
Converted to monashebstemplates

---
title: "Forecasting the old age dependency ratio to determine the best pension age"
author: "Rob J Hyndman, Yijun Zeng and Han Lin Shang"
date: "`r format(Sys.time(), '%-d %B %Y')`"
fontsize: 11pt
papersize: a4
fontfamily: kpfonts
bibliography: pensionage.bib
biblio-style: authoryear-comp
subparagraph: true
toc: false
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(demography)
library(addb)
library(ggplot2)
source("functions-v4.R")
```

\begin{abstract}
We forecast the old-aged dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old aged dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population.

We apply these ideas to Australian historical demographic data and the results suggest that the pension age should be increased to 67 by 2021, 70 by 2032 and 75 by 2058 in order to maintain the old aged dependency ratio at the current level, which is 21%.

This approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries.
\end{abstract}

\begin{keywords}
?????
\end{keywords}

\newpage

#Introduction

Australia's ageing population is a result of three factors: (1) significant technological advances in medical care leading to longer life spans; (2) high fertility rates after World War II resulting in the ``baby-boomer'' generation; and (3) a large decline in fertility rates over the past twenty or so years [@Fehr2008].  Consequently, there are many older people leaving the workforce, and fewer workers replacing them. Similar demographic challenges face many other countries. One way of measuring this problem is via the old-age dependency ratio (OADP): the number of people post-retirement to the number of people of working age [@Walker1982]. The OADP has been increasing for many decades, but recently the rate of increase has jumped markedly, and we show that it will continue to increase at a rapid rate if there are no age-related policy changes.

@@

In this paper, we use the old aged dependency ratio as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07].  The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  OADR=\frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure.  Forecasts of population structure have been built on forecasts of age and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age and gender-specific population structure.

@@

```{r changespension, echo=FALSE, fig.cap="Government's adjustments on pension age scheme.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
year <- 2065-2007
proposed <- ts(pmin(70,
         c(rep(65,2017-2008), 65 + cumsum(rep(c(0.5,0),55))))[1:50],
         start=2008)
approved <- pmin(proposed,67)
current <- pmin(proposed,65)
cols <- c(current='#1b9e77',approved='#d95f02',proposed='#7570b3')
ggplot() +
  autolayer(proposed, series='proposed') +
  autolayer(approved, series='approved') +
  autolayer(current, series='current') +
  xlab("Year") + ylab("Age of retirement") +
  scale_color_manual(values=cols, name=" ",
                     breaks=c("current","approved","proposed"),
                     labels=c("Pension age at 65","Approved change","Proposed further change"))
```

@@

In fact, all supportive arguments for increasing the pension age in Australia have only justified the need to increase the pension age. No justification on the pace or degree of the change has been provided. Also, the government has not disclosed the method it used to determine the new pension age. It is therefore not surprising that the adjustment of the pension age has been questioned by the public. A consistent method to determine the appropriate pension age would be more persuasive in general. For example, it is preferred by OECD countries that pension age should be linked to life expectancy [@OECD12]. There is a practical issue as to how to link life expectancy with pension ages given that there is no information on future mortality rates [@PC13]. @CEDA07 [p7] proposed directly linking pension age with life expectancy in the way that ``pension age should increase by approximately 50% of any increase in life expectancy''. However, using a ‘rule of thumb’ approach might not be ideal, as 50% of the change in life expectancy may not be a good adjustment to pension age in the long term. My study suggests an approach, linking the pension age with the the aged dependency ratio, based on forecasts of future mortality and fertility rates. My approach has the advantage that it can be used to justify the change in pension age and directly links the pension age with our ultimate goal of a stable financial burden for the public.

@@

The deterministic method is judgment-based. The Australian Bureau of Statistics formulates assumptions on the basis of demographic trends, in conjunction with advice from individual experts or government departments. The projections are not predictions or forecasts, ``but are simply illustrations of the growth and change in population which would occur if certain assumptions about future levels of fertility, mortality, internal migration and overseas migration were to prevail over the projection period'' [@ABS12, p.2]. In addition, they do not attempt to take into account of non-demographic factors, such as major government policy changes and catastrophes. Therefore, they normally report multiple (usually three) series based on different levels of assumptions of demographic factors to allow for variety of population projections.

@@

The model proposed by @HU07 is as follows:
\begin{align}
y_t(x)&=s_t(x)+\sigma_t(x)\epsilon_{t,x}\\
s_t(x)&=\mu(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)
\end{align}
The first equation describes the smoothing process. $y_t(x)$ is the observed data in year $t$ and age $x$, and $s_t(x)$ is the smoothed function of $y_t(x)$, estimated by constrained weighted penalized regression splines [BHT14]. $\sigma_t(x)$ allows the variance to be a function of time and age; and $\epsilon_{t,x}$ are independent identical distributed standard normal error terms. The second equation has a similar structure to the LC method, except that it incorporates more than one principal component and score. It describes the dynamics of $s_t(x)$ over time. The $\beta_{t,j}$ for every $j$ and $\mu(x)$ are sets of age-specific constants, and $\phi_j(x)$ for every $j$ is a time-varying index. The $e_t(x)$ is the model error, which is the residual function with mean zero and no serial correlation. The $\mu(x)$ is estimated by the average of $s_t(x)$ across years, which measures the general shape of the observed data $y_t(x)$. The estimates of $\phi_j(x)$ and $\beta_{t,j}$ are the $j$th principal component and scores of $s_t(x)-\mu(x)$ respectively. In other words, we take singular value decomposition of $s_t(x)-\mu(x)$, then equation (7) now becomes
\begin{equation}
s_t(x)-\mu(x)=\phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_k(x)\beta_{t,k}
\end{equation}
We chose $J$ in a way that we believe the majority of information for $s_t(x)-\mu(x)$ has been captured by the first $J$ components.

In terms of forecasting, by conditioning on the observed data $\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}\ $ and the set of functional principal components $\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}$, the $h$-step-ahead forecast of log rate $y_{n+h}(x)$ can be acquired by:
\begin{equation}
\hat{y}_{n+h\mid n}(x)=\hat{s}_{n+h\mid n}(x)=E[\hat{s}_{n+h\mid n}(x)\mid \bm{I, B}]=\hat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\hat{\beta}_{n+h\mid n,j}
\end{equation}
where $\hat{\beta}_{n+h\mid n,j}$ denotes the $h$-step-ahead forecast of $\beta_{n+h\mid n,j}$. Because of the time-varying property of $\left\{\beta_{t,j}\right\}$, it controls the dynamics of the process. For every component series $\left\{\beta_{t,j}\right\}$, we fit them with an $ARIMA(p_j,d_j,q_j)$ process, where $p_j,d_j,q_j$ are determined by minimizing the Akaike Information Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process as follows:
\begin{equation}
(1-\phi_1B-\dots-\phi_{p_j}B^{p_j})(1-B)^{d_j}\beta_{t,j}=c+(1+\theta_1B+\dots+\theta_{q_j}B^{q_j})e_t ~\forall j=1,2..,J
\end{equation}
where the parameters are estimated using maximum likelihood estimation.

@@

We can obtain historical mortality rates $m_t(x)$ (also called central death rates) and fertility rates $f_t(x)$ as follows:
\begin{align*}
m_t(x)&=\frac{D_t(x)}{E_t(x)}=\text{age-sex-specific central death rates in calendar year t}\\
f_t(x)&=B_t(x)/E^F_t(x)=\text{age-specific fertility rates in calendar year t}
\end{align*}

@@

###Step 1 {-}

The main procedure for forecasting the population age structure follows @HB08, except that the data is updated, and we incorporate the product-ratio method [@HBY13] to improve the accuracy of the forecast. The @HU07 model, introduced in section 2.3.2, is applied to the mortality rate, fertility rate and net migration separately, based on the assumption that they behave independently. Therefore, to fit the model to historical fertility rates, the $y_t(x)$ in equation (16) denotes the observed historical fertility data. However, for mortality rate by sex, the product-ratio method is used to ensure the forecast of females and males are non-divergent or coherent [@HBY13]. In the product-ratio method, after smoothing the female and male mortality rates, two other series were created: product series (denoted by $p_t(x)$) and ratio series (denoted by $r_t(x)$), which are defined as follows:
\begin{align*}
  p_t(x)=\sqrt{s_{t,M}(x)s_{t,F}(x)},
  r_t(x)=\sqrt{\frac{s_{t,M}(x)}{s_{t,F}(x)}}
\end{align*}
Then, @HU07 was applied to these two new variables instead of male and female mortality rate as follows:
\begin{align*}
  log[p_t(x)]=\mu_p(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)\\
  log[r_t(x)]=\mu_r(x)+\sum_{l=1}^{L}\gamma_{t,l}\psi_l(x)+z_t(x)
\end{align*}
where time series $\left\{\beta_{t,j}\right\}$ are fitted with an ARIMA model while $\left\{\gamma_{t,l}\right\}$ are fitted by stationary ARFIMA model (i.e., $d$ in ARIMA($p,d,q$) constrained to $0<d<\frac{1}{2}$). It is the stationary constraint on $\gamma_{t,l}$ that ensures the forecast of male and female do not diverge.
Then, simulation h-step-ahead forecast values of product variable and ratio variable was performed (denoted by $p_{n+h\mid n}(x)$ and $r_{n+h\mid n}(x)$ respectively) according to equation (18) and (19).
From here, obtaining h-step-ahead forecast of sex-specific coherent mortality rates (denoted $m_{n+h\mid n,M}(x)$ and $m_{n+h\mid n,F}(x)$) was performed using a back-transformation:
\begin{align*}
m_{n+h\mid n,M}(x)&=p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
m_{n+h\mid n,F}(x)&=p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}
The method for simulating $h$-step-ahead values of net migration by sex is similar to the above method used on mortality rates except no log scale is applied to net migration. A stationary constraint was imposed on the ARIMA model for net migration to make sure the difference between sexes did not diverge.

@@

Suppose the pension age for $n+1$ is $a.r$ where $a$ is the integer and and $r$ is the reminder of pension age. Recall that for $h=1$, we can express the simulated population paths as a matrix with age in the row and simulation in the column as follows:
\begin{equation}
\bm{p}_{n+1}=[p_{mi}]=
\left[
\begin{array}{ccc}
 p_{1,1} & \cdots\cdots & p_{1,1000} \\
\vdots & \ddots & \vdots \\
p_{100^+,1} & \cdots\cdots &p_{100^+,1000}
\end{array}
\right]
\end{equation}
When $r=0$, then the simulation of a population aged 15 to pension age is a subset of the above matrix from $15^{th}$ row to $a-1^{th}$ row, which is:
\begin{equation}
\left[
\begin{array}{ccc}
 p_{15,1} & \cdots\cdots & p_{15,1000} \\
\vdots & \ddots & \vdots \\
p_{a-1,1} & \cdots\cdots &p_{a-1,1000}
\end{array}
\right]
\end{equation}
To calculate the total number of working age, we take a column sum of the trimmed matrix (22). The simulations of total number of work age become a vector:
\begin{equation}
\bm{p}^{work}_{n+1}=[\sum_{m=15}^{a-1}p_{m,1},\sum_{m=15}^{a-1}p_{m,2},\dots,\sum_{m=15}^{a-1}p_{m,1000}]
\end{equation}
Using the same method, the simulations of total number of aged pensioners in the population can be shown as:
\begin{equation}
\bm{p}^{aged}_{n+1}=[\sum_{m=a}^{100^+}p_{m,1},\sum_{m=a}^{100^+}p_{m,2},\dots,\sum_{m=a}^{100^+}p_{m,1000}]
\end{equation}

When $r\neq0$, the vector of total number of work age (23) is added, and the vecor of total number of aged population is subtracted with an adjustment:
\begin{equation}
\text{adjustment}=0.r\times100%[p_{a,1},p_{a,2},\dots,p_{a,1000}]
\end{equation}
This adjustment is based on the assumption that the birthday of population is uniformly distributed over the year. Then, $r\times100\%$ of population with a recorded integer age of $a$ at beginning of the year is actually younger than $a.r$, so they are still under the pension age. In other words,$(1-r\times100\%)$ of population with recorded integer age $a$ at beginning of the year is actually older than $a.r$, so they should be counted as the aged population. The simulation vectors of total work age population and total aged population should then be:
\begin{align*}
\bm{p}^{work}_{n+1}&=[p^{work}_{n+1,1},p^{work}_{n+1,2},\dots,p^{work}_{n+1,1000}]\\
&=[\sum_{m=15}^{a-1}p_{m,1},\sum_{m=15}^{a-1}p_{m,2},\dots,\sum_{m=15}^{a-1}p_{m,1000}]+0.r\times100\%[p_{a,1},p_{a,2},\dots,p_{a,1000}]   \\
\bm{p}^{aged}_{n+1}&=[p^{aged}_{n+1,1},p^{aged}_{n+1,2},\dots,p^{aged}_{n+1,1000}]\\
&=[\sum_{m=a}^{100^+}p_{m,1},\sum_{m=a}^{100^+}p_{m,2},\dots,\sum_{m=a}^{100^+}p_{m,1000}]-0.r\times100%[p_{a,1},p_{a,2},\dots,p_{a,1000}]
\end{align*}

After the adjustment, we can generate simulation vectors of the one-step-ahead OADR by dividing every element of $\bm{p}^{work}_{n+1}$ by the corresponding element of $\bm{p}^{aged}_{n+1}$, i.e.:
\begin{align*}
\bm{OADR}_{n+1}&=[OADR_{n+1,1},OADR_{n+1,2},\dots,OADR_{n+1,1000}]\\
&=\left[\frac{p^{work}_{n+1,1}}{p^{aged}_{n+1,1}},\dots,\frac{p^{work}_{n+1,1000}}{p^{aged}_{n+1,1000}}\right]
\end{align*}
The average value of elements of $\bm{OADR}_{n+1}$ is the mean prediction of one-step-ahead OADR (i.e mean forecast of OADR in 2011).

The same algorithm is applied to the simulation matrix of the population age structure to obtain  $h$-step-ahead simulation vectors of OADR for $h=1,2,\dots,H$, where $H=50$ for this study (i.e., a prediction horizon of 2011 to 2060). Simulations of OADR over a 50 year prediction horizon by the matrix is summarised below, with the forecast year in rows and simulation in the columns:
\begin{equation}
\bm{OADR}^{forecast}=[OADR_{h,i}]=
\left[
\begin{array}{ccc}
 OADR_{n+1,1} & \cdots\cdots & OADR_{n+50,1000} \\
\vdots & \ddots & \vdots \\
OADR_{n+50,1} & \cdots\cdots &OADR_{n+50,1000}
\end{array}
\right]
\end{equation}

@@

```{r getdata, cache=TRUE}
############################
## GET ALL THE DATA
############################

# Get population data for 1 Jan 2010
# Store data locally ????????????
#pop2010 <- hmd.pop("AUS",username,password,"Australia")
#pop2010 <- extract.ages(pop2010, age=0:100)

# Get mortality data
# Store data locally ????????????
#aus.mort <- hmd.mx("AUS",username,password,"Australia")
#aus.mort <- extract.years(extract.ages(aus.mort, 0:100,
#                    combine.upper=FALSE), 
#                    year=1950:max(aus.mort$year))

##### FERTILITY

# Get fertility data:
# Need to install addb package from http://robjhyndman.com/software/addb/
#library(addb) # Contains fertility data for 1921-2006
#aus.fert <- aus.fertility

>>>>

---
title: "Forecasting the old age dependency ratio to determine the best pension age"
author:
- familyname: Hyndman
  othernames: Rob J
  address: Department of Econometrics & Business Statistics\newline Monash University, Clayton VIC 3800, Australia
  email: Rob.Hyndman@monash.edu
  correspondingauthor: true
- familyname: Zeng
  othernames: Yijun
  address: Monash University
- familyname: Shang
  othernames: Hanlin
  address: Australian National University
abstract: "We forecast the old-aged dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old aged dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. We apply these ideas to Australian historical demographic data and the results suggest that the pension age should be increased to 67 by 2021, 70 by 2032 and 75 by 2058 in order to maintain the old aged dependency ratio at the current level, which is 21%. This approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries."
keywords: "blah, blah"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: pensionage.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warning = FALSE, cache = TRUE)
# Make sure you have the latest version of rmarkdown and bookdown
# devtools::install_github("rstudio/rmarkdown")
# devtools::install_github("rstudio/bookdown")
# And the latest versions of my packages
# devtools::install_github("robjhyndman/forecast")
# devtools::install_github("robjhyndman/demography")
# devtools::install_github("robjhyndman/addb")
library(demography)
library(addb)
library(ggplot2)
source("functions-v4.R")
```

# Introduction

Australia's ageing population is a result of three factors: (1) significant technological advances in medical care leading to longer life spans; (2) high fertility rates after World War II resulting in the "baby-boomer" generation; and (3) a large decline in fertility rates over the past twenty or so years [@Fehr2008].  Consequently, there are many older people leaving the workforce, and fewer workers replacing them. Similar demographic challenges face many other countries. One way of measuring this problem is via the old-age dependency ratio (OADP): the number of people post-retirement to the number of people of working age [@Walker1982]. The OADP has been increasing for many decades, but recently the rate of increase has jumped markedly, and we show that it will continue to increase at a rapid rate if there are no age-related policy changes.

@@

In this paper, we use the old aged dependency ratio as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs [@ARW07; @IGR02; @IGR07].  The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age [@ARW07]. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method [@HU07] and the product-ratio method [@HBY13] to forecast a population age structure.  Forecasts of population structure have been built on forecasts of age and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age and gender-specific population structure.

@@

```{r changespension, echo=FALSE, fig.cap="Government's adjustments on pension age scheme.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
year <- 2065 - 2007
proposed <- ts(pmin(
  70,
  c(rep(65, 2017 - 2008), 65 + cumsum(rep(c(0.5, 0), 55)))
)[1:50],
start = 2008
)
approved <- pmin(proposed, 67)
current <- pmin(proposed, 65)
cols <- c(current = "#1b9e77", approved = "#d95f02", proposed = "#7570b3")
ggplot() +
  forecast::autolayer(proposed, series = "proposed") +
  forecast::autolayer(approved, series = "approved") +
  forecast::autolayer(current, series = "current") +
  xlab("Year") + ylab("Age of retirement") +
  scale_color_manual(
    values = cols, name = " ",
    breaks = c("current", "approved", "proposed"),
    labels = c("Pension age at 65", "Approved change", "Proposed further change")
  )
```

@@

In fact, all supportive arguments for increasing the pension age in Australia have only justified the need to increase the pension age. No justification on the pace or degree of the change has been provided. Also, the government has not disclosed the method it used to determine the new pension age. It is therefore not surprising that the adjustment of the pension age has been questioned by the public. A consistent method to determine the appropriate pension age would be more persuasive in general. For example, it is preferred by OECD countries that pension age should be linked to life expectancy [@OECD12]. There is a practical issue as to how to link life expectancy with pension ages given that there is no information on future mortality rates [@PC13]. @CEDA07 [p7] proposed directly linking pension age with life expectancy in the way that "pension age should increase by approximately 50% of any increase in life expectancy". However, using a "rule of thumb" approach might not be ideal, as 50% of the change in life expectancy may not be a good adjustment to pension age in the long term. My study suggests an approach, linking the pension age with the the aged dependency ratio, based on forecasts of future mortality and fertility rates. My approach has the advantage that it can be used to justify the change in pension age and directly links the pension age with our ultimate goal of a stable financial burden for the public.

@@

The deterministic method is judgment-based. The Australian Bureau of Statistics formulates assumptions on the basis of demographic trends, in conjunction with advice from individual experts or government departments. The projections are not predictions or forecasts, "but are simply illustrations of the growth and change in population which would occur if certain assumptions about future levels of fertility, mortality, internal migration and overseas migration were to prevail over the projection period" [@ABS12, p.2]. In addition, they do not attempt to take into account of non-demographic factors, such as major government policy changes and catastrophes. Therefore, they normally report multiple (usually three) series based on different levels of assumptions of demographic factors to allow for variety of population projections.

@@

The model proposed by @HU07 is as follows:
\begin{align}
  y_t(x) & = s_t(x) + \sigma_t(x)\epsilon_{t,x}\\
  s_t(x) & = \mu(x) + \sum_{j=1}^{J}\beta_{t,j}\phi_j(x) + e_t(x)
\end{align}
The first equation describes the smoothing process. $y_t(x)$ is the observed data in year $t$ and age $x$, and $s_t(x)$ is the smoothed function of $y_t(x)$, estimated by constrained weighted penalized regression splines [BHT14]. $\sigma_t(x)$ allows the variance to be a function of time and age; and $\epsilon_{t,x}$ are independent identical distributed standard normal error terms. The second equation has a similar structure to the LC method, except that it incorporates more than one principal component and score. It describes the dynamics of $s_t(x)$ over time. The $\beta_{t,j}$ for every $j$ and $\mu(x)$ are sets of age-specific constants, and $\phi_j(x)$ for every $j$ is a time-varying index. The $e_t(x)$ is the model error, which is the residual function with mean zero and no serial correlation. The $\mu(x)$ is estimated by the average of $s_t(x)$ across years, which measures the general shape of the observed data $y_t(x)$. The estimates of $\phi_j(x)$ and $\beta_{t,j}$ are the $j$th principal component and scores of $s_t(x)-\mu(x)$ respectively. In other words, we take singular value decomposition of $s_t(x)-\mu(x)$, then equation (7) now becomes
\begin{equation}
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_k(x)\beta_{t,k}
\end{equation}
We chose $J$ in a way that we believe the majority of information for $s_t(x)-\mu(x)$ has been captured by the first $J$ components.

In terms of forecasting, by conditioning on the observed data $\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}$ and the set of functional principal components $\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}$, the $h$-step-ahead forecast of log rate $y_{n+h}(x)$ can be acquired by:
\begin{equation}
  \hat{y}_{n+h\mid n}(x)=\hat{s}_{n+h\mid n}(x) = \E[\hat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \hat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\hat{\beta}_{n+h\mid n,j}
\end{equation}
where $\hat{\beta}_{n+h\mid n,j}$ denotes the $h$-step-ahead forecast of $\beta_{n+h\mid n,j}$. Because of the time-varying property of $\left\{\beta_{t,j}\right\}$, it controls the dynamics of the process. For every component series $\left\{\beta_{t,j}\right\}$, we fit them with an ARIMA$(p_j,d_j,q_j)$ process, where $p_j,d_j,q_j$ are determined by minimizing the Akaike Information Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process as follows:
\begin{equation}
  (1-\phi_1B-\dots-\phi_{p_j}B^{p_j})(1-B)^{d_j}\beta_{t,j}=c+(1+\theta_1B+\dots+\theta_{q_j}B^{q_j})e_t ~\forall j=1,2,\dots,J,
\end{equation}
where the parameters are estimated using maximum likelihood estimation.

@@

We can obtain historical mortality rates $m_t(x)$ (also called central death rates) and fertility rates $f_t(x)$ as follows:
\begin{align*}
  m_t(x) & = \frac{D_t(x)}{E_t(x)}=\text{age-sex-specific central death rates in calendar year t}\\
  f_t(x) & = B_t(x)/E^F_t(x)=\text{age-specific fertility rates in calendar year t}
\end{align*}

@@

### Step 1 {-}

The main procedure for forecasting the population age structure follows @HB08, except that the data is updated, and we incorporate the product-ratio method [@HBY13] to improve the accuracy of the forecast. The @HU07 model, introduced in section 2.3.2, is applied to the mortality rate, fertility rate and net migration separately, based on the assumption that they behave independently. Therefore, to fit the model to historical fertility rates, the $y_t(x)$ in equation (16) denotes the observed historical fertility data. However, for mortality rate by sex, the product-ratio method is used to ensure the forecast of females and males are non-divergent or coherent [@HBY13]. In the product-ratio method, after smoothing the female and male mortality rates, two other series were created: product series (denoted by $p_t(x)$) and ratio series (denoted by $r_t(x)$), which are defined as follows:
\begin{align*}
  p_t(x)=\sqrt{s_{t,M}(x)s_{t,F}(x)},
  r_t(x)=\sqrt{\frac{s_{t,M}(x)}{s_{t,F}(x)}}
\end{align*}
Then, @HU07 was applied to these two new variables instead of male and female mortality rate as follows:
\begin{align*}
  log[p_t(x)]=\mu_p(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)\\
  log[r_t(x)]=\mu_r(x)+\sum_{l=1}^{L}\gamma_{t,l}\psi_l(x)+z_t(x)
\end{align*}
where time series $\left\{\beta_{t,j}\right\}$ are fitted with an ARIMA model while $\left\{\gamma_{t,l}\right\}$ are fitted by stationary ARFIMA model (i.e., $d$ in ARIMA($p,d,q$) constrained to $0<d<\frac{1}{2}$). It is the stationary constraint on $\gamma_{t,l}$ that ensures the forecast of male and female do not diverge. Then, simulation h-step-ahead forecast values of product variable and ratio variable was performed (denoted by $p_{n+h\mid n}(x)$ and $r_{n+h\mid n}(x)$ respectively) according to equation (18) and (19). From here, obtaining h-step-ahead forecast of sex-specific coherent mortality rates (denoted $m_{n+h\mid n,M}(x)$ and $m_{n+h\mid n,F}(x)$) was performed using a back-transformation:
\begin{align*}
  m_{n+h\mid n,M}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n,F}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}
The method for simulating $h$-step-ahead values of net migration by sex is similar to the above method used on mortality rates except no log scale is applied to net migration. A stationary constraint was imposed on the ARIMA model for net migration to make sure the difference between sexes did not diverge.

@@

Suppose the pension age for $n+1$ is $a.r$ where $a$ is the integer and and $r$ is the reminder of pension age. Recall that for $h=1$, we can express the simulated population paths as a matrix with age in the row and simulation in the column as follows:
\begin{equation}
  \bm{p}_{n+1}=[p_{mi}] =
    \begin{bmatrix}
       p_{1,1} & \cdots\cdots & p_{1,1000} \\
      \vdots & \ddots & \vdots \\
      p_{100^+,1} & \cdots\cdots &p_{100^+,1000}
      \end{bmatrix}.
\end{equation}
When $r=0$, then the simulation of a population aged 15 to pension age is a subset of the above matrix from $15^{th}$ row to $a-1^{th}$ row, which is:
\begin{equation}
  \begin{bmatrix}
    p_{15,1} & \cdots\cdots & p_{15,1000} \\
    \vdots & \ddots & \vdots \\
    p_{a-1,1} & \cdots\cdots &p_{a-1,1000}
  \end{bmatrix}.
\end{equation}
To calculate the total number of working age, we take a column sum of the trimmed matrix (22). The simulations of total number of work age become a vector:
\begin{equation}
  \bm{p}^{work}_{n+1} = [\sum_{m=15}^{a-1}p_{m,1},\sum_{m=15}^{a-1}p_{m,2},\dots,\sum_{m=15}^{a-1}p_{m,1000}]
\end{equation}
Using the same method, the simulations of total number of aged pensioners in the population can be shown as:
\begin{equation}
  \bm{p}^{aged}_{n+1}=[\sum_{m=a}^{100^+}p_{m,1},\sum_{m=a}^{100^+}p_{m,2},\dots,\sum_{m=a}^{100^+}p_{m,1000}]
\end{equation}

When $r\neq0$, the vector of total number of work age (23) is added, and the vecor of total number of aged population is subtracted with an adjustment:
\begin{equation}
  \text{adjustment}=0.r\times100%[p_{a,1},p_{a,2},\dots,p_{a,1000}]
\end{equation}
This adjustment is based on the assumption that the birthday of population is uniformly distributed over the year. Then, $r\times100\%$ of population with a recorded integer age of $a$ at beginning of the year is actually younger than $a.r$, so they are still under the pension age. In other words,$(1-r\times100\%)$ of population with recorded integer age $a$ at beginning of the year is actually older than $a.r$, so they should be counted as the aged population. The simulation vectors of total work age population and total aged population should then be:
\begin{align*}
  \bm{p}^{work}_{n+1}
    & = [p^{work}_{n+1,1},p^{work}_{n+1,2},\dots,p^{work}_{n+1,1000}]\\
    & = [\sum_{m=15}^{a-1}p_{m,1},\sum_{m=15}^{a-1}p_{m,2},\dots,\sum_{m=15}^{a-1}p_{m,1000}]+0.r\times100\%[p_{a,1},p_{a,2},\dots,p_{a,1000}]   \\
  \bm{p}^{aged}_{n+1}
    & = [p^{aged}_{n+1,1},p^{aged}_{n+1,2},\dots,p^{aged}_{n+1,1000}]\\
    & = [\sum_{m=a}^{100^+}p_{m,1},\sum_{m=a}^{100^+}p_{m,2},\dots,\sum_{m=a}^{100^+}p_{m,1000}]-0.r\times100%[p_{a,1},p_{a,2},\dots,p_{a,1000}]
\end{align*}

After the adjustment, we can generate simulation vectors of the one-step-ahead OADR by dividing every element of $\bm{p}^{work}_{n+1}$ by the corresponding element of $\bm{p}^{aged}_{n+1}$, i.e.:
\begin{align*}
  \bm{OADR}_{n+1} & = [OADR_{n+1,1},OADR_{n+1,2},\dots,OADR_{n+1,1000}]\\
                  & = \left[\frac{p^{work}_{n+1,1}}{p^{aged}_{n+1,1}},\dots,\frac{p^{work}_{n+1,1000}}{p^{aged}_{n+1,1000}}\right]
\end{align*}
The average value of elements of $\bm{OADR}_{n+1}$ is the mean prediction of one-step-ahead OADR (i.e mean forecast of OADR in 2011).

The same algorithm is applied to the simulation matrix of the population age structure to obtain  $h$-step-ahead simulation vectors of OADR for $h=1,2,\dots,H$, where $H=50$ for this study (i.e., a prediction horizon of 2011 to 2060). Simulations of OADR over a 50 year prediction horizon by the matrix is summarised below, with the forecast year in rows and simulation in the columns:
\begin{equation}
  \bm{OADR}^{forecast}=[OADR_{h,i}] =
  \begin{bmatrix}
    OADR_{n+1,1} & \cdots\cdots & OADR_{n+50,1000} \\
    \vdots & \ddots & \vdots \\
    OADR_{n+50,1} & \cdots\cdots &OADR_{n+50,1000}
  \end{bmatrix}.
\end{equation}

@@

```{r getdata, include=FALSE}
############################
## GET ALL THE DATA
############################

# HMD credentials
username <- "rob.hyndman@buseco.monash.edu.au"
password <- "Q234.,mnb"

# Get population data for 1 Jan 2015
auspop <- hmd.pop("AUS", username, password, "Australia")
auspop <- extract.ages(auspop, age = 0:100)

# Get mortality data
aus.mort <- hmd.mx("AUS", username, password, "Australia")
aus.mort <- extract.years(extract.ages(aus.mort, 0:100,
  combine.upper = FALSE
),
year = 1950:max(aus.mort$year)
)
